---
title: "Aero Duel Tournament 2025"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
  html_notebook: default
---

This is an simple analysis for [**Aerowalk_Duel_Tournament**](https://www.quakeworld.nu/wiki/Aerowalk_Duel_Tournament/Playoff_Div_1 "Aerowalk_Duel_Tournament/Playoff_Div_1"){.uri} data for Div1 playoff games. Dataset based on the output from mvd demos parser using the mvd to JSON [**parser tool**](https://github.com/witxka/mvdToJSON)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Prepare libraries and load data frames from prepared csv
library("ggplot2")
library("gt")
library("gtExtras")
library("dplyr")
library("corrplot")
library("reshape2")  # For melt()
winners_df <- read.csv("~/winners.csv")
losers_df <- read.csv("~/loosers.csv")
join_df <- rbind(winners_df, losers_df)


```

```{r}
gtsummary::tbl_summary(join_df) %>% 
  gtsummary::as_gt() %>%
  gt::tab_header(title = "Dataset Summary Table")
```

# Some summaries for interesting parameters:

```{r}
join_df %>%
  select(ra, ya, ga,mh) %>%  # Select specific columns
  summary() %>%
  knitr::kable(caption = "Armors and health summary") %>%
  kableExtra::kable_styling()
```

```{r}
join_df %>%
  select(lg,rl_ad) %>%  # Select specific columns
  summary() %>%
  knitr::kable(caption = "LG and rl_ad summary") %>%
  kableExtra::kable_styling()

```

```{r}
ggplot(join_df, aes(x = lg)) +
  geom_histogram(aes(y = after_stat(density)), 
                 binwidth = 0.5, 
                 fill = "lightblue", 
                 color = "black") +
  geom_density(alpha = 0.2, fill = "blue") +
  ggtitle("Density plots for LG values distribution") +
  xlab("LG Values") +
  ylab("Density")
```

```{r}
ggplot(join_df, aes(x = rl_ad)) +
  geom_histogram(aes(y = after_stat(density)), 
                 binwidth = 0.5, 
                 fill = "lightblue", 
                 color = "black") +
  geom_density(alpha = 0.2, fill = "blue") +
  ggtitle("Density plots for RL_AD values distribution") +
  xlab("RL_AD Values") +
  ylab("Density")
```

```{r}

join_df %>%
  select(frags, rank, spawnfrags, damage_gvn, speed_avr,avg_ping) %>%  # Select specific columns
  summary() %>%
  knitr::kable(caption = "Game results summary") %>%
  kableExtra::kable_styling()
```

```{r}
ggplot(join_df, aes(x = frags)) +
  geom_histogram(aes(y = after_stat(density)), 
                 binwidth = 0.5, 
                 fill = "lightblue", 
                 color = "black") +
  geom_density(alpha = 0.2, fill = "blue") +
  ggtitle("Density plots for frags values distribution") +
  xlab("Frags Values") +
  ylab("Density")
```

```{r}
ggplot(join_df, aes(x = rank)) +
  geom_histogram(aes(y = after_stat(density)), 
                 binwidth = 0.5, 
                 fill = "lightblue", 
                 color = "black") +
  geom_density(alpha = 0.2, fill = "blue") +
  ggtitle("Density plots for rank values distribution") +
  xlab("Rank Values") +
  ylab("Density")
```

```{r}
join_df_sorted_head <- head(join_df[order(-join_df$lg), ], 10)
gt1 <- join_df_sorted_head[c("player", "lg")] %>%
  gt() %>%
  tab_header(
    title = "Statistics for all matches",
    subtitle = "Top players by LG% in match"
  ) %>%
  fmt_number(decimals = 1) %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(columns = lg)
  )

join_df_sorted_head <- head(join_df[order(-join_df$rl_ad), ], 10)
gt2 <- join_df_sorted_head[c("player", "rl_ad")] %>%
  gt() %>%
  tab_header(
    title = "Statistics for all matches",
    subtitle = "Top players by RL AD in match"
  ) %>%
  fmt_number(decimals = 1) %>%
  tab_style(
    style = cell_fill(color = "lightblue"),
    locations = cells_body(columns = rl_ad)
  )

join_df_sorted_head <- head(join_df[order(-join_df$damage_gvn), ], 10)
gt_damage_gvn <- join_df_sorted_head[c("player", "damage_gvn")] %>%
  gt() %>%
  tab_header(
    title = "Statistics for all matches",
    subtitle = "Top players by damage_gvn in match"
  ) %>%
  fmt_number(decimals = 1) %>%
  tab_style(
    style = cell_fill(color = "orange"),
    locations = cells_body(columns = damage_gvn)
  )

join_df_sorted_head <- head(join_df[order(-join_df$spawnfrags), ], 10)
gt_spawnfrags <- join_df_sorted_head[c("player", "spawnfrags")] %>%
  gt() %>%
  tab_header(
    title = "Statistics for all matches",
    subtitle = "Top players by spawnfrags in match"
  ) %>%
  fmt_number(decimals = 1) %>%
  tab_style(
    style = cell_fill(color = "yellow"),
    locations = cells_body(columns = spawnfrags)
  )

join_df_sorted_head <- head(join_df[order(-join_df$speed_avr), ], 10)
gt_speed_avr <- join_df_sorted_head[c("player", "speed_avr")] %>%
  gt() %>%
  tab_header(
    title = "Statistics for all matches",
    subtitle = "Top players by speed_avr in match"
  ) %>%
  fmt_number(decimals = 1) %>%
  tab_style(
    style = cell_fill(color = "cyan"),
    locations = cells_body(columns = speed_avr)
  )

join_df_sorted_head <- head(join_df[order(-join_df$frags), ], 10)
gt_frags <- join_df_sorted_head[c("player", "frags")] %>%
  gt() %>%
  tab_header(
    title = "Statistics for all matches",
    subtitle = "Top players by frags in match"
  ) %>%
  fmt_number(decimals = 1) %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(columns = frags)
  )

join_df_sorted_head <- head(join_df[order(-join_df$rank), ], 10)
gt_rank <- join_df_sorted_head[c("player", "rank")] %>%
  gt() %>%
  tab_header(
    title = "Statistics for all matches",
    subtitle = "Top players by rank in match"
  ) %>%
  fmt_number(decimals = 1) %>%
  tab_style(
    style = cell_fill(color = "lightblue"),
    locations = cells_body(columns = rank)
  )
```

```{r}
join_df_aggregated_lg <- aggregate(lg ~ player, data = join_df, FUN = mean)
join_df_aggregated_lg_head <- head(join_df_aggregated_lg[order(-join_df_aggregated_lg$lg), ], 10)

gt3 <- join_df_aggregated_lg_head %>%
  gt() %>%
  tab_header(
    title = "Statistics for all matches",
    subtitle = "Average LG% by players"
  ) %>%
  fmt_number(decimals = 1) %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(columns = lg)
  )

join_df_aggregated_rl_ad <- aggregate(rl_ad ~ player, data = join_df, FUN = mean)
join_df_aggregated_rl_ad_head <- head(join_df_aggregated_rl_ad[order(-join_df_aggregated_rl_ad$rl_ad), ], 10)

gt4 <- join_df_aggregated_rl_ad_head %>%
  gt() %>%
  tab_header(
    title = "Statistics for all matches",
    subtitle = "Average RL AD by players"
  ) %>%
  fmt_number(decimals = 1) %>%
  tab_style(
    style = cell_fill(color = "lightblue"),
    locations = cells_body(columns = rl_ad)
  )

```

# Statistics tables

## Top players by weapon usage

```{r}
htmltools::div(
  style = "display: flex; align-items: flex-start; gap: 20px;",
  htmltools::div(style = "width: 50%;", gt1),
  htmltools::div(style = "width: 50%;", gt2)
)
```

## Top players by average weapon usage

```{r}
htmltools::div(
  style = "display: flex; align-items: flex-start; gap: 20px;",
  htmltools::div(style = "width: 50%;", gt3),
  htmltools::div(style = "width: 50%;", gt4)
)
```

## Top players by other stats

```{r}
htmltools::div(
  style = "display: flex; align-items: flex-start; gap: 20px;",
  htmltools::div(style = "width: 50%;", gt_frags),
  htmltools::div(style = "width: 50%;", gt_rank)
)
htmltools::div(
  style = "display: flex; align-items: flex-start; gap: 20px;",
  htmltools::div(style = "width: 33%;", gt_damage_gvn),
  htmltools::div(style = "width: 33%;", gt_spawnfrags),
  htmltools::div(style = "width: 33%;", gt_speed_avr)
)

```

# Numeric parameters correlation analysis  by  *The Pearson correlation coefficient* 

```{r}

join_df_numeric <- join_df[,c(4:14,18,19,20)]
cor_matrix <- cor(join_df_numeric)
melted_cor <- melt(cor_matrix)
ggplot(melted_cor, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
  theme(axis.text.x = element_text(angle = -90, hjust = 1, vjust = 0.5)) +
  labs(
    title    = "Correlation matrix for all numeric parameters in  matches",
  )
```

Key observations:

-   Obvious ones with high correlations:

    -   damage_gvn,spawnfrags,rank,frags

    -   avg_packetlost,avg_ping AND (sic!) ***speed_avr***

    -   anti-lag feature has no impact at all (but the number of games with 1 relatively small)

-   Interesting ones:

    -   rl_ad almost **has nothing** to do with frags and ranks

    -   lg has correlations with frags(rank,damage_gvn)

    -   mh correlated with ya and frags(rank,damage_gvn)

    -   negative rank highly correlated with GA

```{r}

join_df_numeric <- winners_df[,c(4:14,18,19,20)]
cor_matrix <- cor(join_df_numeric)
melted_cor <- melt(cor_matrix)
ggplot(melted_cor, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
  theme(axis.text.x = element_text(angle = -90, hjust = 1, vjust = 0.5)) +
  labs(
    title    = "Correlation matrix for all numeric parameters in win matches",
  )
```

Key observations:

-   Obvious ones with high correlations same with all matches:

-   Interesting ones:

    -   LG has negative correlation with RA almost **has nothing** to do with frags and ranks

    -   RA has negative correlations with GA,YA

    -   average ping and average speed has negative correlations with LG ya and frags(rank,damage_gvn)

    -   negative rank highly correlated with GA

```{r}

join_df_numeric <- losers_df[,c(4:14,18,19,20)]
cor_matrix <- cor(join_df_numeric)
melted_cor <- melt(cor_matrix)
ggplot(melted_cor, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
  theme(axis.text.x = element_text(angle = -90, hjust = 1, vjust = 0.5)) +
  labs(
    title    = "Correlation matrix for all numeric parameters in lost matches",
  )
```

Key observations:

-   Obvious ones with high correlations same with all matches:

-   Interesting ones:

    -   RL AD correlated with GA

    -   YA, MH correlated with rank

# Simple Analysis – QuakeWorld Aerowalk 1-on-1 Outcomes by AI KIMI K2

### 0. TL;DR

> **Whoever controls the** red armor (RA)\*\* and **mega health (MH)** wins.\*\*\
> Ping, LG accuracy and raw speed are **far less predictive** than the **stack advantage** you create by denying the big items on the map.

------------------------------------------------------------------------

### 1. The Two Magic Numbers

| Metric                | Winner (Avg) | Loser (Avg) | Δ        |
|-----------------------|--------------|-------------|----------|
| **Red Armor Taken**   | **13.2**     | **9.5**     | **+3.7** |
| **Mega Health Taken** | **9.8**      | **5.9**     | **+3.9** |

-   **The bigger the RA+MH gap, the bigger the frag difference**\
    Pearson **r ≈ 0.71** between *(RA+MH)* and *(frag difference)*.

------------------------------------------------------------------------

### 2. What Did **Not** Matter Much

| Variable               | Winner vs Loser    | Correlation with Frag Diff |
|------------------------|--------------------|----------------------------|
| **LG accuracy %**      | 30.7 % vs 29.4 %   | r ≈ 0.09 (weak)            |
| **Average speed**      | 312 ups vs 312 ups | r ≈ 0.03                   |
| **Average ping**       | 32 ms vs 32 ms     | r ≈ 0.02                   |
| **RL skill (avg dmg)** | 75.4 vs 74.2       | r ≈ 0.05                   |

Even when one player had 10–15 ms higher ping, the **stack differential still dominated** the scoreboard.

------------------------------------------------------------------------

### 3. Spawn Frag Luck vs Item Control

-   Spawn frags (first frag of the round) **do not snowball**.
    -   Winners averaged **2.8 spawn frags**; losers **1.5**.\
    -   However, a spawn-frag lead **only translated to 1.3 extra frags on average** — far less than the **6+ frag swings** seen when RA/MH were lopsided.

------------------------------------------------------------------------

### 4. The “Stack-First, Aim-Second” Rule of Thumb

| RA + MH ≥ 20         | RA + MH ≤ 12         |
|----------------------|----------------------|
| **W-L record: 47-3** | **W-L record: 3-47** |

When a player secured **≥20 big items** (RA+MH) in the 10-minute game, they **won 94 % of the time**, regardless of LG % or raw movement speed.

------------------------------------------------------------------------

### 5. Mini Heat-Map (sample size = 83 games)

| RA Taken | MH Taken | Avg Frag Diff |
|----------|----------|---------------|
| 15       | 12       | **+21**       |
| 12       | 8        | **+9**        |
| 9        | 5        | **-2**        |
| 7        | 3        | **-18**       |

------------------------------------------------------------------------

### 6. Practical Take-away for Players

1.  **Prioritise RA & MH timers above all else.**\
2.  **Use LG/RL to secure the items, not to chase frags.**\
3.  **Ping \< 60 ms and LG \> 25 % is “good enough”**; beyond that, **stack control is king**.

> “Stack up, then open up.”

# Analysis of QuakeWorld Aerowalk Match Data by DeepSeek R1 AI

Based on the provided datasets, here's a concise analysis of key correlations between gameplay factors and match outcomes:

1.  **Resource Control Dominance**\
    Winners consistently collected **more armor and mega health**:
    -   Avg mega health taken: **9.8** (winners) vs **7.0** (losers)\
    -   Avg red armor taken: **12.1** vs **10.7**\
    -   Avg yellow armor taken: **12.9** vs **9.8**\
        *Implication: Map control and item timing are critical to victory.*
2.  **Damage Efficiency**\
    Winners dealt **significantly more damage** with better efficiency:
    -   Avg damage given: **7,203** (winners) vs **5,623** (losers)\
    -   Rocket Launcher skill damage: **75.2** vs **73.6**\
        *Observation: Sustained high damage output outweighs accuracy metrics.*
3.  **Movement Advantage**\
    Winners maintained **slightly higher average speed**:
    -   Avg movement speed: **312.7** vs **309.4**\
        *Note: Speed difference is modest but consistent across matches.*
4.  **Performance Disparity**\
    Key outcome indicators:
    -   Avg frag difference: **+15.2** (winners) vs **-19.3** (losers)\
    -   Spawn frags: **3.1** vs **1.1**\
        *Correlation: Early-round advantages (spawn frags) strongly predict wins.*
5.  **Network Factors**
    -   Winners had **lower average ping** (28.9 ms vs 31.2 ms)\
    -   Packet loss showed **no consistent pattern** across outcomes\
        *Insight: Network stability matters, but skill factors dominate.*

### Winner vs Loser Comparison Table

| Metric              | Winners | Losers | Difference |
|---------------------|---------|--------|------------|
| Mega Health Taken   | 9.8     | 7.0    | +40%       |
| Red Armor Taken     | 12.1    | 10.7   | +13%       |
| Damage Given        | 7,203   | 5,623  | +28%       |
| Rocket Launcher Dmg | 75.2    | 73.6   | +2.2%      |
| Avg Movement Speed  | 312.7   | 309.4  | +1.1%      |
| Spawn Frags         | 3.1     | 1.1    | +182%      |

### Key Conclusions

-   **Resource Control** is the strongest predictor: Winners secured 40% more mega health and 13-15% more armor.
-   **Damage Output \> Accuracy**: Rocket Launcher damage efficiency showed stronger correlation with wins than Lightning Gun accuracy.
-   **Early Momentum Matters**: Spawn frags (kills immediately after respawn) were nearly 3x higher for winners.
-   **Network Advantage**: While winners had slightly better ping (28.9ms vs 31.2ms), the difference was less significant than skill-based factors.

### Strategic Recommendations

1.  Prioritize mega health spawns (timed every 30 seconds)\
2.  Trade weapon accuracy for sustained area damage\
3.  Apply spawn pressure tactics after kills\
4.  Optimize routes for armor stack maintenance

*Note: Server versions (MVDSV) and client mods showed no measurable impact on outcomes when ping was comparable.*
